### 2020-09-19 backup/end/main.tex
integrate older

\documentclass[12pt]{article}

\usepackage{amsmath, amsthm, amssymb}
\usepackage{amssymb}
\usepackage{bbm}

\usepackage[hidelinks]{hyperref}
\usepackage{cleveref}
\usepackage{url}

%%% For figures
\usepackage{graphicx}
\DeclareGraphicsExtensions{.pdf,.png,.jpg}

%%% For tables
\usepackage[table]{xcolor}

%specifying table and figure packages
\usepackage{caption}

\usepackage[T1]{fontenc}
\usepackage[all]{xy}
\usepackage[inline]{enumitem}

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{remark}{Remark}
\newtheorem{example}{Example}
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}

\newcommand{\mattia}[1]{\textcolor{cyan}{#1}}
\newcommand{\pietro}[1]{\textcolor{teal}{#1}}
\newcommand{\Ban}{{\mathbf{Ban}}}
\newcommand{\Top}{{\mathbf{Top}}}
\newcommand{\forget}{{\textnormal{forget}}}
\newcommand{\free}{{\textnormal{free}}}
\newcommand{\pt}{{\textnormal{pt}}}
\newcommand{\Hom}{{\textnormal{Hom}}}
\newcommand{\End}{{\textnormal{End}}}
\newcommand{\Fun}{{\textnormal{Fun}}}
\newcommand{\Aut}{{\textnormal{Aut}}}
\newcommand{\Obj}{{\textnormal{Obj}}}
\newcommand{\id}{{\textnormal{id}}}
\newcommand{\Morph}{{\textnormal{Morph}}}
\newcommand{\Set}{{\mathbf{Set}}}
\newcommand{\Cat}{{\mathbf{C}}}
\newcommand{\DCat}{{\mathbf{D}}}
\newcommand{\ICat}{{\mathbf{I}}}
\newcommand{\JCat}{{\mathbf{J}}}
\newcommand{\LCat}{{\mathbf{L}}}
\newcommand{\range}[2]{{\{{#1}, \dots,{#2}\}}}
\newcommand{\anon}{{\,\mbox{-}\,}}

\crefname{diagram}{diag.}{diags.}
\Crefname{diagram}{Diagram}{Diagrams}
\creflabelformat{diagram}{#2(#1)#3}

\title{Comonadic neural network architectures}
\author{
    Pietro Vertechi \and Mattia G. Bergomi
}
\date{}
\begin{document}
\maketitle
\begin{abstract}
\end{abstract}

\section{Architecture}

We begin by giving an abstract definition of {\em architecture}. \pietro{Non so se vogliamo questa come definizione, oppure se prendere come definizione le ipotesi di \cref{thm:kan_architecture}.}

\begin{definition}\label{def:architecture}
    An {\em architecture} is a natural transformation from a comonad to an endofunctor. More explicitly, given a category $\DCat$, an {\em architecture} on $\DCat$ is composed of the following data:
    \begin{itemize}
        \item a comonad $T$ on $\DCat$,
        \item an endofunctor $S \colon \DCat \rightarrow \DCat$,
        \item a natural transformation $\eta \colon T \rightarrow S$.
    \end{itemize}
\end{definition}

% An architecture $(\DCat, T, S, \eta)$ is {\em valid} if the following conditions are met:
% \begin{itemize}
%     \item $\DCat$ has finite products,
%     \item for all $(X, +)$ commutative monoid in $\DCat$ \pietro{meglio introdurre la categoria lineare a parte}, and for all
%           \begin{equation*}
%               f \colon SX \rightarrow X,
%           \end{equation*}
%           the composition
%           \begin{equation*}
%               f \circ \eta_X \colon TX \rightarrow X
%           \end{equation*}
%           is a {\em machine}~\cite{2020arXiv200702777V} in the Kleisli category $\DCat_T$.
% \end{itemize}

\pietro{Da discutere il caso continuo (interpretazione comonadica di macchine di Volterra). Probabilmente da fare con enriched category theory.}

This is sufficient to give notions of composability, equivariance, and locality. In the reminder of the section, we will first give a broad class of architectures. Then, we will consider a simpler special case and show that it is sufficient to recover, among other examples, classical recurrent and convolutional neural networks.

Option: drop this and do it for a diagram
\begin{equation*}
    M \rightarrow C_1 \rightrightarrows C_0.
\end{equation*}

\begin{theorem}\label{thm:kan_architecture}
    Let $p\colon \ICat \rightarrow \JCat$ be a functor between small categories. Let $H$ be a subfunctor of the composition
    \begin{equation*}
        \ICat^{op} \times \ICat \xrightarrow{p \times p}
        \JCat^{op} \times \JCat \xrightarrow{\Hom} \Set.
    \end{equation*}
    Let $\Cat$ be a complete category. Then, $p$ and $H$ induce an architecture on $\Cat^\ICat$, which we call a {\em Kan architecture}.
\end{theorem}

\begin{proof}
    $p$ induces a functor $p^* \colon \Cat^\JCat \rightarrow \Cat^\ICat$, given by pre-composition. As $\Cat$ is complete, $p^*$ admits a right adjoint $p_* \colon \Cat^\JCat \rightarrow \Cat^\ICat$, its {\em right Kan extension}~\cite[Chapt.~X]{mac2013categories}. This adjunction $p^* \dashv p_*$ induces a comonad
    \begin{equation*}
        T := p^*p_* \colon \Cat^\ICat \rightarrow \Cat^\ICat.
    \end{equation*}
    Given $i \in \Obj(\ICat)$, $T(i)$ can be described explicitly using the {\em end} formula,~\cite[Eq.~X.4.3]{mac2013categories}, which we report here in our notation for a functor $F \colon \ICat \rightarrow \Cat$:
    \begin{equation}\label{eq:endformula}
        p_* F = \int_i F(i)^{\Hom(\anon, p(i))}.
    \end{equation}
    The integral represents an {\em end}~\cite[Sect.~IX.5]{mac2013categories} over the category $\ICat$. From~\cref{eq:endformula} we can deduce that,
    \begin{equation}\label{eq:comonad_endformula}
        T F = \int_i F(i)^{\Hom(p(\anon), p(i))}.
    \end{equation}
    We then define
    \begin{equation}\label{eq:functor_endformula}
        S F := \int_i F(i)^{H(\anon, i)},
    \end{equation}
    where $S F$ is a functor, as $H$ is closed by pre-composition of morphisms in the image of $p$. By functoriality of end, the inclusion $H \hookrightarrow \Hom \circ p \times p$ gives us a natural transformation
    \begin{equation*}
        \eta \colon T \rightarrow S.
    \end{equation*}
\end{proof}

\subsection{Simple Kan architecture}

\Cref{thm:kan_architecture} gives an explicit example of architecture that is still somewhat complex. Let us consider a particular case. Let  $p\colon J \hookrightarrow \JCat$ be the underlying discrete subcategory, having the same objects as $\JCat$, but only identity morphisms. In this scenario, \cref{thm:kan_architecture} simplifies greatly. In particular, $H(i, j)$ can simply be written as $\Hom_\JCat(i, j) \cap M$, for some $M \subseteq \Morph(\JCat)$. Furthermore, ends over discrete categories are just products, so~\cref{eq:comonad_endformula,eq:functor_endformula} become:
\begin{equation}\label{eq:productformula}
    T G = \prod_{j \in J} G(j)^{\Hom_\JCat(\anon, j)}
    \quad\text{ and }\quad
    S G = \prod_{j \in J} G(j)^{Hom_\JCat(\anon, j) \cap M}.
\end{equation}
This result is summarized in the following theorem.

\begin{theorem}\label{thm:discrete_kan_architecture}
    Let $\JCat$ be a small category, and let $J = \Obj(\JCat)$ be its underlying discrete category. Let $M \subseteq \Morph(\JCat)$ be a subset of morphisms in $\JCat$. Let $\Cat$ be a category with small products. Then $\JCat$ and $M$ induce an architecture on $\Cat^J$.
\end{theorem}

\begin{proof}
    The comonad and functor are given by~\cref{eq:productformula}.
\end{proof}

\section{Comonadic machines}

Show how it's possible to obtain a machine from an architecture.

\section{Compactly-generated Hausdorff spaces}

\section{Recurrent and convolutional neural networks}

Monoids can be considered as categories with one object. Let us consider the category corresponding to $\mathbb N$. Let $M$ be the set containing only the generator morphism. Then the above construction recovers recurrent neural networks. Analogously, considering $\mathbb N \times \mathbb N$ we obtain convolutional neural networks, where $M$ denotes the shape of the filter.
\pietro{TODO: mention 2 objects for conv, and strides!}
\pietro{Maybe also do functors, to generalize GENEO? Somehow discuss continuous version?}

\section{Novel architectures}

Here, we explore the scenario where the category $\Cat$ has more than one object.
\pietro{TODO: figure out a good application.}
\pietro{Mention partial symmetries because of non-composability.}

\bibliographystyle{abbrv}
\bibliography{References}

\end{document}

### 2020-09-19 backup/simpleproduct/main.tex
integrate older

\documentclass[12pt]{article}

\usepackage{amsmath, amsthm, amssymb}
\usepackage{amssymb}
\usepackage{bbm}

\usepackage[hidelinks]{hyperref}
\usepackage{cleveref}
\usepackage{url}

%%% For figures
\usepackage{graphicx}
\DeclareGraphicsExtensions{.pdf,.png,.jpg}

%%% For tables
\usepackage[table]{xcolor}

%specifying table and figure packages
\usepackage{caption}

\usepackage[T1]{fontenc}
\usepackage[all]{xy}
\usepackage[inline]{enumitem}

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{remark}{Remark}
\newtheorem{example}{Example}
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}

\newcommand{\mattia}[1]{\textcolor{cyan}{#1}}
\newcommand{\pietro}[1]{\textcolor{teal}{#1}}
\renewcommand{\P}{{\mathcal{P}}}
\newcommand{\Ban}{{\mathbf{Ban}}}
\newcommand{\Top}{{\mathbf{Top}}}
\newcommand{\forget}{{\textnormal{forget}}}
\newcommand{\free}{{\textnormal{free}}}
\newcommand{\pt}{{\textnormal{pt}}}
\newcommand{\Hom}{{\textnormal{Hom}}}
\newcommand{\End}{{\textnormal{End}}}
\newcommand{\Aut}{{\textnormal{Aut}}}
\newcommand{\Obj}{{\textnormal{Obj}}}
\newcommand{\id}{{\textnormal{id}}}
\newcommand{\Morph}{{\textnormal{Morph}}}
\newcommand{\Cat}{{\mathbf{C}}}
\newcommand{\LCat}{{\mathbf{L}}}
\newcommand{\range}[2]{{\{{#1}, \dots,{#2}\}}}
\newcommand{\anon}{{\,\mbox{-}\,}}
\newcommand{\causal}{{\Delta_{\P, \Cat}}}

\crefname{diagram}{diag.}{diags.}
\Crefname{diagram}{Diagram}{Diagrams}
\creflabelformat{diagram}{#2(#1)#3}

\title{
    Causal machines
    }
\author{
    Pietro Vertechi \and Mattia G. Bergomi
}
\date{}
\begin{document}
    \maketitle
    \begin{abstract}
    \end{abstract}

    \section{Causal category}
    Let $\Cat$ be a category with finite products. Let $\P = (P, \preceq)$ be a finite preordered set. Let us consider the functor category $[P, \Cat]$, where $P$ is seen as a discrete category. We recall, and slightly adapt, a definition of comonad originally given in~\cite{uustaluComonadicNotionsComputation2008}.
    \begin{definition}\label{def:causalcomonad}
        The {\em causal comonad} $D\colon [P, \Cat] \rightarrow [P, \Cat]$ is given by
        \begin{equation*}
            (DA)_p = \prod_{q \preceq p} A_q.
        \end{equation*}
        $D$ is equipped with a counit $\epsilon\colon D \rightarrow \id$ given by the projection $\pi_p$:
        \begin{equation*}
            \epsilon_A \colon \prod_{q \preceq p} A_q \rightarrow A_p.
        \end{equation*}
        It is also equipped with a comultiplication $\delta\colon D \rightarrow DD$:
        \begin{equation*}
            \delta_A \colon \prod_{q \preceq p} A_q \rightarrow \prod_{q \preceq p} \left(\prod_{r \preceq q} A_r\right),
        \end{equation*}
        induced by the projections $\prod_{q \preceq p} A_q \rightarrow \prod_{r \preceq q} A_r$.
    \end{definition}
    \begin{definition}\label{def:causalcategory}
        The {\em causal category} $\causal$ is the Kleisli category associated to the causal comonad $D$.
    \end{definition}

    While \cref{def:causalcomonad,def:causalcategory} seem abstract, it is possible to describe the causal category $\causal$ explicitly.
    Objects in $\causal$ are collections of objects of $\Cat$ indexed by $P$, that is $\{A_p\}_{p\in P}$. Let us consider two such objects $\{A_p\}_{p\in P}, \{B_p\}_{p\in P} \in \Obj(\causal)$. Then,
    \begin{equation*}
        \Hom_\causal\left(\{A_p\}_{p\in P}, \{B_p\}_{p\in P}\right) =
        \prod_{p \in P} \Hom_\Cat\left(\prod_{q \preceq p} A_q, B_p\right).
    \end{equation*}

    \section{Stream comonad in Cartesian closed categories}
    Here, we recall the notion of {\em stream comonad}~\cite{uustaluComonadicNotionsComputation2008} in a general Cartesian closed category.
    Let $\Cat$ be a Cartesian closed category. $\Cat$ is a monoidal category, with monoidal structure given by the Cartesian product. Let us consider the functor category
    \begin{equation*}
        \EndCat.
    \end{equation*}
    $\EndCat$ is also a monoidal category, with monoidal structure given by the composition of functors. There is a functor
    \begin{align*}
        \Cat^{op} &\rightarrow \EndCat \\
        M &\mapsto [M, -],
    \end{align*}
    which preserves the monoidal structure, as $[M \times N, -]$ is naturally isomorphic to $[M, [N, -]]$, for $M, N \in \Obj(\Cat)$.
    As a consequence, whenever $M \in \Obj(\Cat)$ is a monoid, $[M, -]$ is a comonad.
    \begin{definition}\label{def:streamcomonad}\cite[Sect.~3.1]{uustaluComonadicNotionsComputation2008}
        Let $M$ be a monoid in a Cartesian closed category $\Cat$.
        The {\em stream comonad} $D_M\colon \Cat \rightarrow \Cat$ is given by
        \begin{equation*}
            D_M A = [M, A]
        \end{equation*}
        The {\em stream category} $\Stream$ is the Kleisli category associated to the stream comonad $D_M$.
    \end{definition}

    In the following sections, we will focus on $\StreamN$ and $\StreamR$, which correspond to discrete-time and continuous-time recurrent architectures.

        % \section{Categorical preliminaries}

    % Let $\Cat$ be a category with small products, and let $\JCat$ be a small category.
    % The aim of this section is to describe concretely a specific functor
    % \begin{equation*}
    %    \Cat \rightarrow \Fun(\JCat, \Cat).
    % \end{equation*}
    % Given $j\in\Obj(\JCat)$, we can consider the functor
    % \begin{align*}
    %     \Phi_j &\colon \Cat \rightarrow \FunCat\\
    %     \Phi_j(X) &:= i \mapsto X ^ {\Hom_\JCat(i, j)}.
    % \end{align*}
    % Multiplying all these functors together, we obtain
    % \begin{equation*}
    %     \Phi := \prod_{j \in \Obj(\JCat)} \Phi_j \colon \Cat \rightarrow \FunCat.
    % \end{equation*}
    % Let $X, Y\in\Obj(\Cat)$. It is possible to describe $\Hom_{\FunCat}(\Phi(X), \Phi(Y))$ explicitly.

    % \begin{lemma}\label{lm:yoneda_prod}
    %     Let $F \in \FunCat$, $j \in \JCat$, and $Y \in \Obj(\Cat)$. Then,
    %     \begin{equation*}
    %         \Hom_{\FunCat}(F, \Phi_j(Y)) \simeq \Hom_\Cat(F(j), Y).
    %     \end{equation*} 
    % \end{lemma}
    % \begin{proof}
    %     \begin{align*}
    %         \Hom_{\FunCat}(F, \Phi_j(Y))
    %         &\simeq \Hom_{\FunCat}\left(F, Y ^ {\Hom_\JCat(\anon, j)}\right) \\
    %         &\simeq \Hom_{\PreSheaves}\left(\Hom_\JCat(\anon, j), \Hom_\Cat(F(\anon), Y)\right) \\
    %         &\simeq \Hom_\Cat(F(j), Y),
    %     \end{align*}
    %     where the last isomorphism is the Yoneda lemma.
    % \end{proof}

    % \begin{lemma}\label{lm:yoneda_all}
    %     Let $F \in \FunCat$, $j \in \JCat$, and $Y \in \Obj(\Cat)$. Then,
    %     \begin{equation*}
    %         \Hom_{\FunCat}(F, \Phi(Y)) \simeq \prod_{j\in \Obj(\JCat)}\Hom_\Cat(F(j), Y).
    %     \end{equation*} 
    % \end{lemma}
    % \begin{proof}
    %     By~\cref{lm:yoneda_prod}.
    % \end{proof}

    % \begin{proposition}\label{prop:morphisms}
    %     Let $X, Y \in \Obj(\Cat)$. Then         
    %     \begin{equation*}
    %         \Hom_{\FunCat}(\Phi(X), \Phi(Y)) \simeq
    %         \prod_{j\in \Obj(\JCat)}\Hom_\Cat\left(X^{out(j)}, Y\right),
    %     \end{equation*}
    %     where $out(j)$ is the set of morphisms whose domain is $j$.
    % \end{proposition}
    % \begin{proof}
    %     By~\cref{lm:yoneda_all}.
    % \end{proof}

    % \begin{proposition}\label{prop:product}
    %     $\Phi$ preserves small products.
    % \end{proposition}
    % \begin{proof}
    %     \pietro{TODO}
    % \end{proof}

    % \begin{remark}
    %     Discuss comonad approach from \cite{Ahman_2016}.
    % \end{remark}

    % \subsection{Recurrent neural networks}

    % Let us consider the category $\mathbf N$, with one object whose endomorphisms are given by the monoid $\mathbb N$. Then the above construction recovers recurrent neural networks.

    \section{Categorical preliminaries}

    Let $\Cat$ be a cocomplete Cartesian closed category, and let $\PCat$ be a small category.
    The aim of this section is to describe how a presheaf $P$ on $\PCat$ induces an adjunction
    \begin{align*}
        L_P &\colon \Fun(\PCat, \Cat) \rightarrow \Cat,\\
        R_P &\colon \Cat \rightarrow \Fun(\PCat, \Cat).
    \end{align*}
    Let $\PreSheaves$ denote the category of presheaves on $\PCat$. Let us define $R$ as follows.
    \begin{align*}
        R \colon \PreSheaves^{op} &\rightarrow \Fun(\Cat, \Fun(\PCat, \Cat)),\\
        R_P &:= X \mapsto X^{P(\anon)}.
    \end{align*}
    Let us consider:
    \begin{align*}
        L\colon \PCat &\rightarrow \Fun(\Fun(\PCat, \Cat), \Cat)\\
        L_p &:= X \mapsto X(p).
    \end{align*}
    As $\Cat$ has small colimits, $L$ extends uniquely (up to natural isomorphism) to a colimit-preserving functor
    \begin{equation*}
        L\colon \PreSheaves \rightarrow \Fun(\Fun(\PCat, \Cat), \Cat).
    \end{equation*}

    \begin{proposition}\label{prop:adjunction}
        For all $P \in \PreSheaves$, $L_P$ is the left adjoint of $R_P$.
    \end{proposition}
    \begin{proof}
        Let $X \in \Obj(\FunCat)$, $Y \in \Obj(\Cat)$. We wish to prove that
        \begin{equation*}
            \Hom_{\FunCat}(X, R_P(Y)) \simeq \Hom_\Cat(L_P, Y).
        \end{equation*}
        Thanks to the density theorem~\cite[Chapt.~3.7]{mac2013categories}, we can write $P$ as a colimit of representable presheaves:
        \begin{equation*}
            P \simeq \varinjlim \Hom(\anon, p).
        \end{equation*}
        Then,
        \begin{align*}
            \Hom_{\FunCat}(X, R_P(Y))
            &\simeq \Hom_{\FunCat}\left(X, Y ^ {P(\anon)}\right) \\
            &\simeq \Hom_{\PreSheaves}\left(P(\anon), \Hom_\Cat(X(\anon), Y)\right) \\
            &\simeq \Hom_{\PreSheaves}\left(\varinjlim \Hom(\anon, p), \Hom_\Cat(X(\anon), Y)\right) \\
            &\simeq \varprojlim\Hom_{\PreSheaves}\left(\Hom(\anon, p), \Hom_\Cat(X(\anon), Y)\right) \\
            &\simeq \varprojlim\Hom_\Cat(X(p), Y)\\
            &\simeq \Hom_\Cat(\varinjlim X(p), Y)\\
            &\simeq \Hom_\Cat(L_P(X), Y),
        \end{align*}
        where $\Hom_{\PreSheaves}\left(\Hom(\anon, p), \Hom_\Cat(X(\anon), Y)\right) \simeq \Hom_\Cat(X(p), Y)$ follows from the Yoneda lemma.
    \end{proof}

    By~\cref{prop:adjunction}, for any presheaf $P$, the functors $L_P, R_P$ induce a comonad $L_P\circ R_P$ on $\Cat$, which we denote $T_P$. Mostly, we will use the presheaf
    \begin{equation*}
        I = \coprod_{p \in \Obj(\PCat)} \Hom(\anon, p).
    \end{equation*}
    \begin{remark}
        In the case of the presheaf $I$, the comonad $T_I$ is equivalent to a construction derived in the context of {\em directed containers} in~\cite{Ahman_2016}.
    \end{remark}

    \begin{proposition}\label{prop:morphisms}
        Let $X, Y \in \Obj(\Cat)$. Then         
        \begin{equation*}
            \Hom_\Cat(T_I(X), Y) \simeq
            \prod_{p\in \Obj(\PCat)}
            \Hom_\Cat\left(\prod_{q \in \Obj(\PCat)} X^{\Hom(p, q)}, Y\right).
        \end{equation*}
    \end{proposition}
    \begin{proof}
        By~\cref{prop:adjunction}.
    \end{proof}

    \subsection{Recurrent neural networks}

    Let us consider the category $\mathbf N$, with one object whose endomorphisms are given by the monoid $\mathbb N$. Then the above construction recovers recurrent neural networks.

    %% Old simpler version

    Let $\JCat$ be a small category, and let $J = \Obj(\JCat)$ be its underlying set of objects. We can consider $J$ as a category, whose only morphisms are identities. Let $p \colon J \rightarrow \JCat$ be the inclusion of $J$ in $\JCat$. We are interested in extending functors $J \rightarrow \Cat$ to functors $\JCat \rightarrow \Cat$, using a standard construction, {\em right Kan extensions}~\cite[Chapt.~X]{mac2013categories}. For the inclusion $J \rightarrow \JCat$, the explicit formula given in~\cite[Sect.~3.7]{borceux1994handbook}, which requires $\Cat$ to be complete, simplifies in this scenario. In particular, we only require $\Cat$ to have small products.

\begin{proposition}\label{prop:adjunction}
    The functor $p \colon J \rightarrow \JCat$ induces an adjunction
    \begin{align*}
        p^*\colon \Cat^\JCat & \rightarrow \Cat^J
                             &
                             & \dashv
                             &
        p_*\colon \Cat^J     & \rightarrow \Cat^\JCat                         \\
        F                    & \mapsto F\circ p
                             &
                             &
                             &
        G                    & \mapsto \prod_{j \in J} G(j)^{\Hom(\anon, j)}.
    \end{align*}
    That is to say, given functors $F\colon \JCat \rightarrow \Cat$ and $G \colon J \rightarrow \Cat$, there is a natural isomorphism
    \begin{equation*}
        \Hom(p^*F, G) \simeq \Hom(F, p_*G).
    \end{equation*}
\end{proposition}

\begin{proof}
    Thanks to Yoneda's lemma, for all $j \in J$,
    \begin{equation*}
        \Hom(F(j), G(j)) \simeq \Hom(\Hom(\anon, j), \Hom(F(\anon), G(j))).
    \end{equation*}
    As a consequence,
    \begin{align*}
        \Hom(p^* F, G)
         & \simeq \prod_{j\in J} \Hom(F(j), G(j))                            \\
         & \simeq \prod_{j \in J} \Hom(\Hom(\anon, j), \Hom(F(\anon), G(j))) \\
         & \simeq \prod_{j \in J} \Hom(F, G(j)^{\Hom(\anon, j)})             \\
         & \simeq \Hom(F, p_* G),
    \end{align*}
    where all isomorphisms are natural.
\end{proof}

By~\cref{prop:adjunction}, we can define a comonad $T = p^*p_* \colon \Cat^J \rightarrow \Cat^J$. The Kleisli category associated to $T$, denoted $\Cat^J_T$, can be described explicitly. Given $X, Y\in \Obj(\Cat^J)$, we have
\begin{equation*}
    \Hom(T X, Y) \simeq
    \prod_{i \in J} \Hom\left(\prod_{j \in J} X(j)^{\Hom(i, j)}, Y(i)\right).
\end{equation*}

\begin{remark}
    The comonad $T$ is remindful of constructions derived in the context of {\em directed containers} in~\cite{Ahman_2016,uustalu_comonadic_2008}.
\end{remark}

\begin{definition}
    Given a subset of morphisms $M \subseteq \Morph(\JCat)$, we can consider a functor $T_M$, given by
    \begin{align*}
        T_M \colon \Cat^J & \rightarrow \Cat^J                                    \\
        X                 & \mapsto \prod_{j \in J} X(j)^{\Hom(\anon, j) \cap M}.
    \end{align*}
    We denote $\mu$ the obvious natural transformations $T \rightarrow T_M$.
    We say that a map $f\colon TX \rightarrow Y$ is $M$-local if it factors via
    \begin{equation*}
        \mu_X \colon TX \rightarrow T_M X.
    \end{equation*}
\end{definition}

\bibliographystyle{abbrv}
\bibliography{References}

\end{document}

### 2020-09-20 src/volterra.jl
volterra poc

@@ -1,0 +1,46 @@
+using FastGaussQuadrature, LinearAlgebra
+
+function radauII(n)
+    x, w = gaussradau(n)
+    x̂ = reverse((1 .- x) / 2)
+    ŵ = reverse(w / 2)
+    return x̂, ŵ
+end
+
+# idea: use block methods for Kernel case!
+function volterra(K, G, g, mesh, n)
+    cj, wj = radauII(n)
+    is = ntuple(identity, length(mesh) - 1)
+    results = foldl(is, init=()) do res, i
+        pts = @. mesh[i] + cj * (mesh[i+1] - mesh[i])
+        acc = g.(pts)
+        for j in 1:(i-1)
+            pts0 = @. mesh[j] + cj * (mesh[j+1] - mesh[j])
+            wts0 = wj * (mesh[i+1] - mesh[i])
+            mat = K.(pts', pts0) .* wts0
+            acc += mat * res[j]
+        end
+        return (res..., G.(acc))
+    end
+    return last(results)
+end
+
+## Example
+
+function K(t, s)
+    t > s .+ 0.1
+end
+
+mesh = 0:0.1:1
+G(x) = tanh(x)
+g(x) = x
+n = 4
+
+volterra(K, G, g, mesh, n)
+
+using BenchmarkTools
+@benchmark volterra($K, $G, $g, $mesh, $n)
+using ProfileView
+ProfileView.@profview for i in 1:100
+    volterra(K, G, g, mesh, n)
+end

### 2020-10-02 backup/separate/main.tex
massive simplification

\documentclass[12pt]{article}

\usepackage{amsmath, amsthm, amssymb}
\usepackage{amssymb}
\usepackage{bbm}

\usepackage[hidelinks]{hyperref}
\usepackage{cleveref}
\usepackage{url}

%%% For figures
\usepackage{graphicx}
\DeclareGraphicsExtensions{.pdf,.png,.jpg}

%%% For tables
\usepackage[table]{xcolor}

%specifying table and figure packages
\usepackage{caption}

\usepackage[T1]{fontenc}
\usepackage[all]{xy}
\usepackage[inline]{enumitem}

\usepackage{tikz-cd}

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{remark}{Remark}
\newtheorem{example}{Example}
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}

\newcommand{\mattia}[1]{\textcolor{cyan}{#1}}
\newcommand{\pietro}[1]{\textcolor{teal}{#1}}
\newcommand{\N}{{\mathbb{N}}}
\newcommand{\Ban}{{\mathbf{Ban}}}
\newcommand{\Top}{{\mathbf{Top}}}
\newcommand{\pt}{{\textnormal{pt}}}
\newcommand{\Hom}{{\textnormal{Hom}}}
\newcommand{\End}{{\textnormal{End}}}
\newcommand{\Fun}{{\textnormal{Fun}}}
\newcommand{\Aut}{{\textnormal{Aut}}}
\newcommand{\Obj}{{\textnormal{Obj}}}
\newcommand{\id}{{\textnormal{id}}}
\newcommand{\Morph}{{\textnormal{Morph}}}
\newcommand{\Set}{{\mathbf{Set}}}
\newcommand{\Cat}{{\mathbf{C}}}
\newcommand{\DCat}{{\mathbf{D}}}
\newcommand{\FCoalg}{{F\textnormal{-coalg}}}
\newcommand{\stCoalg}{{s_!t^*\textnormal{-coalg}}}
\newcommand{\range}[2]{{\{{#1}, \dots,{#2}\}}}
\newcommand{\anon}{{\,\mbox{-}\,}}

\crefname{diagram}{diag.}{diags.}
\Crefname{diagram}{Diagram}{Diagrams}
\creflabelformat{diagram}{#2(#1)#3}

\title{Weight sharing via comonads}
\author{
    Pietro Vertechi \and Mattia G. Bergomi
}
\date{}
\begin{document}
\maketitle
\begin{abstract}
\end{abstract}

\section{Introduction}

\pietro{Qui o dopo.}
One of the key reasons for the success of neural networks is the notion of {\em weight sharing}. In its simplest form, one can consider, among maps $[M, X] \rightarrow [M, X]$, those that originate from maps $[M, X] \rightarrow X$. Indeed, given
\begin{equation*}
    {\cal F} \colon [M, X] \rightarrow X,
\end{equation*}
one can consider the map
\begin{align*}
    \widehat{\cal F} \colon [M, X]       & \rightarrow [M, X]              \\
    \left(\widehat{\cal F}\phi\right)(m) & = {\cal F}(a \mapsto \phi(ma)).
\end{align*}

This simple notion of weight sharing does not seem sufficient to cover complex neural networks, made up of several nodes, each with their own symmetries, which interact with each other in non-trivial ways.

We aim to show that the simple notion of weight sharing defined above is sufficient to recover, among other examples, classical recurrent and convolutional neural networks. However, to do so, we will need to generalize not the weight sharing strategy, but rather the underlying category on which such strategy takes place.

\section{Categorical preliminaries}

\begin{definition}\label{def:nilpotent_functor}
    A functor $F$ is {\em nilpotent} if there exists $n \in \N$ such that the composed functor $F^n$ maps all objects to the terminal object.
\end{definition}

Graph-like structures are a source of examples of nilpotent functors.

\begin{proposition}\label{prop:graph_architecture}
    Let $(E, V)$ be a directed acyclic graph. Let $\DCat$ be a category with finite products. The functor
    \begin{align*}
        F \colon \DCat^V & \rightarrow \DCat^V                  \\
        (F X)(v)         & = \prod_{u \, | \, (u, v)\in E} X(u)
    \end{align*}
    is nilpotent.
\end{proposition}

\begin{proof}
    Let $n$ be such that there are no paths of length $n$ in $(V, E)$. Then $F^n X$ is the terminal object.
\end{proof}

\Cref{prop:graph_architecture} has a continuous analogue in locally Cartesian closed categories, which will be our focus.

\begin{definition}\label{def:categorical_architecture}
    Let $\DCat$ be a locally Cartesian closed category. An internal quiver
    \begin{equation*}
        s, t\colon E \rightrightarrows V
    \end{equation*}
    {\em has bounded depth} if there exists $n \in \N$ such that the iterated pullback
    \begin{equation}\label{eq:nilpotent_left_adjoint}
        \underbrace{E \times_V E \times_V \dots \times_V E}_{n \text{ factors}}
    \end{equation}
    is initial in $\DCat/V$.
\end{definition}

\begin{proposition}\label{prop:categorical_architecture}
    Let $\DCat$ be a locally Cartesian closed category. Let $s, t\colon E \rightrightarrows V$ be an quiver of bounded depth. Then \pietro{TODO: definisci * e ! (base change functors).}
    \begin{equation*}
        s_!t^*\colon \DCat/V \rightarrow \DCat/V
    \end{equation*}
    is a nilpotent endofunctor.
\end{proposition}

\begin{proof}
    The functor $s_!t^*$ is the right adjoint of the functor  $t_*s^*\colon \DCat/V \rightarrow \DCat/V$. By~\cref{eq:nilpotent_left_adjoint}, $(t_*s^*)^n$ maps all objects to the initial object, so its adjoint $(s_!t^*)^n$ maps all objects to the terminal object.
\end{proof}

\subsection{The category of coalgebras}

Let $\Cat$ be a Cartesian closed category with finite limits, and let $F\colon \Cat \rightarrow \Cat$ be a left exact (i.e., preserves finite limits) nilpotent endofunctor. We can consider $\FCoalg$, the category of coalgebras over $F$ \pietro{add ref}. Objects are simply maps
\begin{equation*}
    X \rightarrow F X,
\end{equation*}
where $X \in \Obj(\Cat)$. Morphisms are given by commutative diagrams
\begin{equation*}
    \begin{tikzcd}
        X \arrow[rightarrow]{d} \arrow[rightarrow]{r}{\phi}
        & Y \arrow[rightarrow]{d} \\
        FX \arrow[rightarrow]{r}{F \phi}
        & FY
    \end{tikzcd}
\end{equation*}
As $F$ is nilpotent and preserves finite products, it is straightforward to verify that the forgetful functor
\begin{equation*}
    U \colon \FCoalg \rightarrow \Cat
\end{equation*}
has a right adjoint $\Phi$, called the cofree functor, given by
\begin{equation*}
    \Phi(X) = X \times F X \times \dots \times F^{n-1} X \rightarrow F X \times \dots \times F^{n-1} X.
\end{equation*}

This adjunction is comonadic \pietro{explain! cite Beck's theorem!}, and $\FCoalg$ is the Eilenberg-Moore category $\Cat^T$, where $T = U\Phi$. As $T$ is left exact (product of left exact), $\FCoalg \simeq \Cat^T$ is Cartesian closed and it has finite limits, which are preserved by both $U$ and $\Phi$~\cite[Sect.~5.50]{wyler1991lecture}.

\begin{proposition}\label{prop:cofree_exponentiating}
    Let $F\colon \Cat \rightarrow \Cat$ be a left exact, nilpotent endofunctor. Then the exponential of cofree coalgebras is again cofree.
    More explicitly, let $\Phi\colon \Cat \rightarrow \FCoalg$ denote the cofree functor. Then, there exists a natural isomorphism
    \begin{equation*}
        [B, \Phi X] \simeq \Phi[UB, X].
    \end{equation*}
\end{proposition}

\begin{proof}
    By a straightforward computation:
    \begin{align*}
        \Hom_{\FCoalg}(A, [B, \Phi X])
         & \simeq \Hom_{\FCoalg}(A\times B, \Phi X)     \\
         & \simeq \Hom_{\Cat}(U(A \times B), X)  \\
         & \simeq \Hom_{\Cat}(UA \times UB, X)   \\
         & \simeq \Hom_{\Cat}(UA, [UB, X])       \\
         & \simeq \Hom_{\FCoalg}(A, \Phi[UB, X]).
    \end{align*}
    By Yoneda's lemma, this induces a natural isomorphism
    \begin{equation*}
        [B, \Phi X] \simeq \Phi[UB, X].
    \end{equation*}
\end{proof}

\section{Machines}

\begin{definition}\label{def:machine}
    Let $\Cat$ be a category, equipped with a terminal object $\pt$. Let $X \in \Obj(\Cat)$. A map
    \begin{equation*}
        \varrho \colon X \rightarrow X
    \end{equation*}
    is a {\em machine} if the induced transformation
    \begin{equation*}
        \Hom_\Cat(\pt, X) \rightarrow \Hom_\Cat(\pt, X)
    \end{equation*}
    has a unique fixed point $S_\varrho$, which we call the {\em stable state} of $\varrho$.
\end{definition}

\pietro{Fare qualche esempio, tipo contrazioni.}

\begin{definition}\label{def:parametric_machine}
    Let $\Cat$ be a category with finite products. A {\em parametric machine} is simply a machine on the Kleisli category for the comonad $P \times \anon$, for some $P \in \Obj(\Cat)$. More explicitly, a map
    \begin{equation*}
        \varrho \colon P \times X \rightarrow X
    \end{equation*}
    is a {\em machine} if there exists a unique map $S_\varrho$ such that the following diagram commutes
    \begin{equation}
        \begin{tikzcd}[column sep=large]\label[diagram]{diag:machine_condition}
            P \arrow[rightarrow]{r}{\left(\id, S_\varrho\right)}
            \arrow[rightarrow,swap]{dr}{S_\varrho}
            & P \times X \arrow[rightarrow]{d}{\varrho} \\
            & X
        \end{tikzcd}
    \end{equation}
\end{definition}

\Cref{diag:machine_condition} is analogous to the machine condition in~\cite{2020arXiv200702777V} in the case when $X$ is an Abelian group and $\varrho\colon P \times X \rightarrow X$ is the sum of a map $P \rightarrow X$ and a map $X \rightarrow X$. This approach differs from the one taken in~\cite{2020arXiv200702777V}, in that here the object $P$ represents both input space and parameters. In what follows, we will show how to construct examples of {\em machines} using {\em nilpotent functors}.

% \begin{definition}\label{def:nilpotent_architecture}
%     A {\em nilpotent architecture} is a left exact, nilpotent endofunctor $F$ on a finitely complete, Cartesian closed category $\Cat$.
% \end{definition}

\begin{theorem}\label{thm:nilpotent_architecture}
    Let $F\colon \Cat \rightarrow \Cat$ be a left exact, nilpotent endofunctor. Let $\Phi(X)$ be a cofree coalgebra in $\FCoalg$. Then a map $F X \rightarrow X$ induces a machine $\Phi X \rightarrow \Phi X$.
\end{theorem}

\begin{proof}
    \pietro{TODO, da fare anche con parametric machines e $P \times F X \rightarrow X$, dove $P$ è una $F$ coalgebra con un'azione di $M$.}
\end{proof}

\begin{definition}\label{def:architecture}
    An {\em architecture} is a quiver of bounded depth $s,t \colon E \rightrightarrows V$ in a locally Cartesian closed category $\DCat$, together with a monoid object $M$ in $\stCoalg$.
\end{definition}

This is sufficient to give notions of composability, weight sharing, and locality.

\begin{theorem}\label{thm:weight_sharing}
    A nilpotent architecture $M, s, t$ induces a comonad on $\DCat/V$, given by
    \begin{equation*}
        X \mapsto U\Phi[UM, X].
    \end{equation*}
\end{theorem}

\begin{proof}
    \pietro{Free forget composed with $[M, \anon]$ in $\stCoalg$.}
\end{proof}

\pietro{Forse si possono riunificare le due nozioni!}

\section{Recurrent and convolutional neural networks}

Monoids can be considered as categories with one object. Let us consider the category corresponding to $\N$. Then the above construction recovers recurrent neural networks. Analogously, considering $\N \times \N$ we obtain convolutional neural networks \pietro{discuss shape of filter?}

\section{Novel architectures}

\bibliographystyle{abbrv}
\bibliography{References}

\end{document}

### 2020-10-08 backup/presimpler/main.tex
before volterra

--- 2020-10-02 backup/separate/main.tex
@@
 \newcommand{\DCat}{{\mathbf{D}}}
+\newcommand{\JCat}{{\mathcal{J}}}
 \newcommand{\FCoalg}{{F\textnormal{-coalg}}}
@@
 
-\pietro{Qui o dopo.}
 One of the key reasons for the success of neural networks is the notion of {\em weight sharing}.
@@
 
-We aim to show that the simple notion of weight sharing defined above is sufficient to recover, among other examples, classical recurrent and convolutional neural networks.
-However, to do so, we will need to generalize not the weight sharing strategy, but rather the underlying category on which such strategy takes place.
+We aim to show that the a similar strategy, based on a different comonad, is sufficient to recover, among other examples, classical recurrent and convolutional neural networks, incorporating at the same time equivariance and locality.
 
 \section{Categorical preliminaries}
+
+\begin{definition}\label{def:cartesian_category}
+    We say that a category is {\em Cartesian} if it has finite products. We say that a functor between Cartesian categories is {\em Cartesian} if it preserves finite products.
+\end{definition}
+
+\pietro{TODO definisci anche locally Cartesian closed!}
+
+\begin{proposition}\label{prop:internal_category_comonad}
+    Let $\DCat$ be a locally Cartesian closed category. Then, there is a well defined notion of category {\em internal} to $\DCat$~\cite{mac2013categories}.
+    Let $\JCat = s, t \colon C_1 \rightrightarrows C_0$ be such an internal category. $\JCat$ induces a limit-preserving comonad
+    \begin{equation*}
+        t_!s^*\colon \DCat/C_0 \rightrightarrows \DCat/C_0.
+    \end{equation*}
+    \pietro{TODO: definisci * e ! (base change functors).}
+\end{proposition}
+
+\begin{proof}
+    The map $s_*t^*\colon \DCat/C_0 \rightrightarrows \DCat/C_0$ is a monad~\cite[Thm.~V.8.2]{Mac_Lane_1994}.
+As $\DCat$ is locally Cartesian closed, $s^*$ has a right adjoint $s_!$, hence we can consider $t_!s^*$.
+As $T = t_!s^*$ is the right adjoint of the monad $s_*t^*$, it is a comonad.
+Being a right adjoint, $T$ preserves limits.
+\end{proof}
 
@@
 
-Graph-like structures are a source of examples of nilpotent functors.
-
-\begin{proposition}\label{prop:graph_architecture}
-    Let $(E, V)$ be a directed acyclic graph. Let $\DCat$ be a category with finite products. The functor
-    \begin{align*}
-        F \colon \DCat^V & \rightarrow \DCat^V                  \\
-        (F X)(v)         & = \prod_{u \, | \, (u, v)\in E} X(u)
-    \end{align*}
-    is nilpotent.
-\end{proposition}
-
-\begin{proof}
-    Let $n$ be such that there are no paths of length $n$ in $(V, E)$. Then $F^n X$ is the terminal object.
-\end{proof}
-
-\Cref{prop:graph_architecture} has a continuous analogue in locally Cartesian closed categories, which will be our focus.
-
-\begin{definition}\label{def:categorical_architecture}
+\begin{definition}\label{def:internal_quiver_bounded_depth}
     Let $\DCat$ be a locally Cartesian closed category. An internal quiver
@@
 
-\begin{proposition}\label{prop:categorical_architecture}
-    Let $\DCat$ be a locally Cartesian closed category. Let $s, t\colon E \rightrightarrows V$ be an quiver of bounded depth. Then \pietro{TODO: definisci * e ! (base change functors).}
+\begin{proposition}\label{prop:internal_quiver_functor}
+    Let $\DCat$ be a locally Cartesian closed category. Let $s, t\colon E \rightrightarrows V$ be an quiver of bounded depth. Then
     \begin{equation*}
@@
     \end{equation*}
-    is a nilpotent endofunctor.
+    is a limit-preserving nilpotent functor.
 \end{proposition}
@@
 \begin{proof}
-    The functor $s_!t^*$ is the right adjoint of the functor  $t_*s^*\colon \DCat/V \rightarrow \DCat/V$.
+    The functor $s_!t^*$ is the right adjoint of the functor  $t_*s^*\colon \DCat/V \rightarrow \DCat/V$, hence it preserves limits and is therefore Cartesian.
 By~\cref{eq:nilpotent_left_adjoint}, $(t_*s^*)^n$ maps all objects to the initial object, so its adjoint $(s_!t^*)^n$ maps all objects to the terminal object.
@@
 
+\section{Equivariant architectures}
+
+Let us now assume that $\DCat$  is a locally Cartesian closed category with a terminal object (and thus all finite limits).
+
+\begin{definition}\label{def:equivariance_category}
+    Let $M$ be a monoid in $\DCat$ and $I \in \Obj$. The {\em equivariance category} $\JCat_{M, I}$ is given by the projection maps
+    \begin{equation*}
+        s, t\colon M \times I \times I \rightrightarrows I,
+    \end{equation*}
+    with composition and identity given by composition and identity in $M$.
+The {\em equivariance comonad} is the comonad induced by this category via~\cref{prop:internal_category_comonad}, that is, $t_!s^*$.
+\end{definition}
+
+We introduce the notion of {\em nilpotent constraint}, which we will need in order to limit equivariance and connectivity. To do so, we need some additional definitions.
+
+\begin{definition}\label{def:nilpotent_architecture}
+    Let $\JCat_{M, I}$ be an equivariance category. A {\em nilpotence constraint} on $\JCat_{M, I}$ is a subobject $S \subseteq M \times I \times I$ such that the internal quiver induced by
+    \begin{equation*}
+        S \hookrightarrow M \times I \times I \rightrightarrows I
+    \end{equation*}
+    has bounded depth, as in~\cref{def:internal_quiver_bounded_depth}.
+\end{definition}
+% Let $\Cat$ be a Cartesian category, and let $T$ be a comonad on $\Cat$.
+A {\em nilpotent constraint} on $T$ is a nilpotent Cartesian endofunctor $F \colon \Cat \rightarrow \Cat$ together with a natural transformation $\eta \colon T \rightarrow F$.
+
+This is sufficient to give a notion of locality.
+
 \subsection{The category of coalgebras}
 
-Let $\Cat$ be a Cartesian closed category with finite limits, and let $F\colon \Cat \rightarrow \Cat$ be a left exact (i.e., preserves finite limits) nilpotent endofunctor.
+Let $\Cat$ be a Cartesian closed category, and let $F\colon \Cat \rightarrow \Cat$ be a nilpotent architecture.
 We can consider $\FCoalg$, the category of coalgebras over $F$ \pietro{add ref}.
@@
 \end{equation*}
+As $F$ is Cartesian, the category $\FCoalg$ is also Cartesian, with products given by
+\begin{equation*}
+    \prod_{i \in I} X_i \rightarrow \prod_{i \in I} F X_i \simeq F \prod_{i \in I} X_i.
+\end{equation*}
 As $F$ is nilpotent and preserves finite products, it is straightforward to verify that the forgetful functor
@@
 
-This adjunction is comonadic \pietro{explain!
-cite Beck's theorem!}, and $\FCoalg$ is the Eilenberg-Moore category $\Cat^T$, where $T = U\Phi$.
-As $T$ is left exact (product of left exact), $\FCoalg \simeq \Cat^T$ is Cartesian closed and it has finite limits, which are preserved by both $U$ and $\Phi$~\cite[Sect.~5.50]{wyler1991lecture}.
-
-\begin{proposition}\label{prop:cofree_exponentiating}
-    Let $F\colon \Cat \rightarrow \Cat$ be a left exact, nilpotent endofunctor. Then the exponential of cofree coalgebras is again cofree.
-    More explicitly, let $\Phi\colon \Cat \rightarrow \FCoalg$ denote the cofree functor. Then, there exists a natural isomorphism
-    \begin{equation*}
-        [B, \Phi X] \simeq \Phi[UB, X].
-    \end{equation*}
-\end{proposition}
-
-\begin{proof}
-    By a straightforward computation:
-    \begin{align*}
-        \Hom_{\FCoalg}(A, [B, \Phi X])
-         & \simeq \Hom_{\FCoalg}(A\times B, \Phi X)     \\
-         & \simeq \Hom_{\Cat}(U(A \times B), X)  \\
-         & \simeq \Hom_{\Cat}(UA \times UB, X)   \\
-         & \simeq \Hom_{\Cat}(UA, [UB, X])       \\
-         & \simeq \Hom_{\FCoalg}(A, \Phi[UB, X]).
-    \end{align*}
-    By Yoneda's lemma, this induces a natural isomorphism
-    \begin{equation*}
-        [B, \Phi X] \simeq \Phi[UB, X].
-    \end{equation*}
-\end{proof}
+\begin{remark}
+    This adjunction is comonadic \pietro{explain! cite Beck's theorem!}, and $\FCoalg$ is the Eilenberg-Moore category $\Cat^T$, where $T = U\Phi$.
+\end{remark}
 
@@
 
-% \begin{definition}\label{def:nilpotent_architecture}
-%     A {\em nilpotent architecture} is a left exact, nilpotent endofunctor $F$ on a finitely complete, Cartesian closed category $\Cat$.
-% \end{definition}
-
 \begin{theorem}\label{thm:nilpotent_architecture}
-    Let $F\colon \Cat \rightarrow \Cat$ be a left exact, nilpotent endofunctor.
-Let $\Phi(X)$ be a cofree coalgebra in $\FCoalg$.
-Then a map $F X \rightarrow X$ induces a machine $\Phi X \rightarrow \Phi X$.
+    Let $F\colon \Cat \rightarrow \Cat$ be a nilpotent endofunctor.
+Let $\Phi X$ be a cofree coalgebra in $\FCoalg$.
+Let $P$ be a $F$ coalgebra.
+Then a map $P \times F X \rightarrow X$ induces a parametric machine $P \times \Phi X \rightarrow \Phi X$.
 \end{theorem}
@@
 \begin{proof}
-    \pietro{TODO, da fare anche con parametric machines e $P \times F X \rightarrow X$, dove $P$ è una $F$ coalgebra con un'azione di $M$.}
+    \pietro{TODO per induzione.}
 \end{proof}
 
-\begin{definition}\label{def:architecture}
-    An {\em architecture} is a quiver of bounded depth $s,t \colon E \rightrightarrows V$ in a locally Cartesian closed category $\DCat$, together with a monoid object $M$ in $\stCoalg$.
-\end{definition}
-
-This is sufficient to give notions of composability, weight sharing, and locality.
-
-\begin{theorem}\label{thm:weight_sharing}
-    A nilpotent architecture $M, s, t$ induces a comonad on $\DCat/V$, given by
-    \begin{equation*}
-        X \mapsto U\Phi[UM, X].
-    \end{equation*}
-\end{theorem}
-
-\begin{proof}
-    \pietro{Free forget composed with $[M, \anon]$ in $\stCoalg$.}
-\end{proof}
-
-\pietro{Forse si possono riunificare le due nozioni!}
+\pietro{La sfida principale qui è trovare un buona architettura $F$ ma anche una buona coalgebra $P$.
+Probabilmente da menzionare da qualche parte che se la categoria di base è buona, $\FCoalg$ ha molte buone proprietà.}
 
@@
 
-Monoids can be considered as categories with one object.
-Let us consider the category corresponding to $\N$.
-Then the above construction recovers recurrent neural networks.
-Analogously, considering $\N \times \N$ we obtain convolutional neural networks \pietro{discuss shape of filter?}
+\pietro{discuss shape of filter?}
 

### 2020-10-08 backup/unify/main.tex
before volterra

--- 2020-10-08 backup/presimpler/main.tex
@@
 \newcommand{\DCat}{{\mathbf{D}}}
-\newcommand{\JCat}{{\mathcal{J}}}
 \newcommand{\FCoalg}{{F\textnormal{-coalg}}}
@@
 
-\section{Categorical preliminaries}
+\section{Architecture}
 
@@
 
-\pietro{TODO definisci anche locally Cartesian closed!}
-
-\begin{proposition}\label{prop:internal_category_comonad}
-    Let $\DCat$ be a locally Cartesian closed category. Then, there is a well defined notion of category {\em internal} to $\DCat$~\cite{mac2013categories}.
-    Let $\JCat = s, t \colon C_1 \rightrightarrows C_0$ be such an internal category. $\JCat$ induces a limit-preserving comonad
-    \begin{equation*}
-        t_!s^*\colon \DCat/C_0 \rightrightarrows \DCat/C_0.
-    \end{equation*}
-    \pietro{TODO: definisci * e ! (base change functors).}
-\end{proposition}
-
-\begin{proof}
-    The map $s_*t^*\colon \DCat/C_0 \rightrightarrows \DCat/C_0$ is a monad~\cite[Thm.~V.8.2]{Mac_Lane_1994}.
-As $\DCat$ is locally Cartesian closed, $s^*$ has a right adjoint $s_!$, hence we can consider $t_!s^*$.
-As $T = t_!s^*$ is the right adjoint of the monad $s_*t^*$, it is a comonad.
-Being a right adjoint, $T$ preserves limits.
-\end{proof}
-
 \begin{definition}\label{def:nilpotent_functor}
@@
 
-\begin{definition}\label{def:internal_quiver_bounded_depth}
+We begin by giving an abstract definition of {\em nilpotent architecture}.
+
+\begin{definition}\label{def:nilpotent_architecture}
+    Let $\Cat$ be a Cartesian category. A {\em nilpotent architecture} is a nilpotent Cartesian endofunctor $F \colon \Cat \rightarrow \Cat$.
+\end{definition}
+
+This is sufficient to give notions of composability, weight sharing, and locality.
+Graph-like structures are a source of examples of nilpotent architecture.
+
+\begin{proposition}\label{prop:graph_architecture}
+    \pietro{Il multigrafo è più interessante, ma tanto è solo un esempio.}
+    Let $(E, V)$ be a directed acyclic graph. Let $\DCat$ be a Cartesian category. The functor
+    \begin{align*}
+        F \colon \DCat^V & \rightarrow \DCat^V                  \\
+        (F X)(v)         & = \prod_{u \, | \, (u, v)\in E} X(u)
+    \end{align*}
+    is a nilpotent architecture.
+\end{proposition}
+
+\begin{proof}
+    It is straightforward to verify that $F$ is Cartesian.
+    Let $n$ be such that there are no paths of length $n$ in $(V, E)$. Then $F^n X$ is the terminal object.
+\end{proof}
+
+\Cref{prop:graph_architecture} has a continuous analogue in locally Cartesian closed categories, which will be our focus.
+
+\begin{definition}\label{def:categorical_architecture}
     Let $\DCat$ be a locally Cartesian closed category. An internal quiver
@@
 
-\begin{proposition}\label{prop:internal_quiver_functor}
-    Let $\DCat$ be a locally Cartesian closed category. Let $s, t\colon E \rightrightarrows V$ be an quiver of bounded depth. Then
+\begin{proposition}\label{prop:categorical_architecture}
+    Let $\DCat$ be a locally Cartesian closed category. Let $s, t\colon E \rightrightarrows V$ be an quiver of bounded depth. Then \pietro{TODO: definisci * e ! (base change functors).}
     \begin{equation*}
@@
     \end{equation*}
-    is a limit-preserving nilpotent functor.
+    is a nilpotent architecture.
 \end{proposition}
@@
 \end{proof}
-
-\section{Equivariant architectures}
-
-Let us now assume that $\DCat$  is a locally Cartesian closed category with a terminal object (and thus all finite limits).
-
-\begin{definition}\label{def:equivariance_category}
-    Let $M$ be a monoid in $\DCat$ and $I \in \Obj$. The {\em equivariance category} $\JCat_{M, I}$ is given by the projection maps
-    \begin{equation*}
-        s, t\colon M \times I \times I \rightrightarrows I,
-    \end{equation*}
-    with composition and identity given by composition and identity in $M$.
-The {\em equivariance comonad} is the comonad induced by this category via~\cref{prop:internal_category_comonad}, that is, $t_!s^*$.
-\end{definition}
-
-We introduce the notion of {\em nilpotent constraint}, which we will need in order to limit equivariance and connectivity. To do so, we need some additional definitions.
-
-\begin{definition}\label{def:nilpotent_architecture}
-    Let $\JCat_{M, I}$ be an equivariance category. A {\em nilpotence constraint} on $\JCat_{M, I}$ is a subobject $S \subseteq M \times I \times I$ such that the internal quiver induced by
-    \begin{equation*}
-        S \hookrightarrow M \times I \times I \rightrightarrows I
-    \end{equation*}
-    has bounded depth, as in~\cref{def:internal_quiver_bounded_depth}.
-\end{definition}
-% Let $\Cat$ be a Cartesian category, and let $T$ be a comonad on $\Cat$.
-A {\em nilpotent constraint} on $T$ is a nilpotent Cartesian endofunctor $F \colon \Cat \rightarrow \Cat$ together with a natural transformation $\eta \colon T \rightarrow F$.
-
-This is sufficient to give a notion of locality.
 
